Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-01-12 E-Hypergraphs-LICS24.tex
Sec 6 figure update

CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
Because extended cospans have a more general notion of interface, DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
First, the existence of \textit{internal} interfaces needs to be accounted for,  and in particular that internal interfaces are not necessarily preserved during rewriting.
The former means that we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching external interfaces.

\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\]
and
\[
    i \xrightarrow{} i'' \xrightarrow{} \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j
\]

We will next use the fact that cospans
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\]
and
\[
    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
\]
endcode the same data and only consider cospans of the form

\[
\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\] to fit the DPO formalism.
We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of $\mathcal{G}$ should be modified to exclude the vertices that were a part of (the image of) $\mathcal{L}$'s strict internal interfaces.
Then, when gluing $\mathcal{R}$ into the created context, we need to introduce strict internal interfaces of $\mathcal{R}$ into this context.
In particular, it means that the internal interfaces of the resulting cospan of e-hypergraphs
\[
    \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k 
\]
not necessarily coincides with the internal interfaces of the initial cospan.
% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
% Note that the outermost interfaces are preserved.

@@

\begin{figure}
    \[
    \scalebox{0.6}{
    \tikzfig{combinatorial_semantics/f_plus_g_new}
    }
    \]
    \caption{$+$ of two morphisms in $\Ecospans$}
    \label{fig:A+B}
\end{figure}

@@

Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms). 
% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
CHRIS: COMMENT THAT WE QUOTIENT TO PHRASE CATEGORICALLY OUR RESULTS. 
\begin{definition}[Quotient by rewrites]  
Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
\[
	f \sim g \quad \text{if} \quad f \Rightarrow^*_{\llbracket \mathcal E \rrbracket} g ~ . 
\]
\end{definition}
In particular,  let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
This quotient in fact turns the interpretation function into a proper functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which can further be quotiented to account for extra equations $\mathcal E$ arising in symmetric monoidal theories,  as below. 
\[
	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
\]

Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.  
%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
Thus we have the following result. 
\begin{proposition}[Soundness]
The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
\end{proposition}
It immediately follows that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is a semilattice-enriched PROP,  and $\llbracket - \rrbracket_{\mathcal E}$ is the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself. 
Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.

CHRIS: NOTE ALSO WEAK NORMAL FORM FOR $\Sigma^+$-TERMS?
The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}. 
\begin{proposition}[Weak normal form]
For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
\[
	f = f_1 + \ldots + f_n
\]
 such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
\end{proposition}

@@

Unfolding the definitions,  we can spell out the previous theorem in terms of DPO-rewriting. 
\begin{corollary}[Completeness of rewriting]
For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
\[
	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^*_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
\]
 \end{corollary}
%
%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
%
%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
%\end{theorem}
%
%\begin{proposition}
%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
%\end{proposition}
%
%\begin{theorem}
%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
%\end{theorem}
CHRIS: CONCLUDING STATEMENT.  COMMENT ABOUT GENERALIZATION TO DIRECTED REWRITES.

@@

\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
\label{th:existence_of_pushouts}
Consider the following span in $\catname{EHyp_{\Sigma}}$
% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
\[\begin{tikzcd}
	Z && X \\
	\\
	Y
	\arrow["f", from=1-1, to=1-3]
	\arrow["g"', from=1-1, to=3-1]
\end{tikzcd}\]
such that
\begin{enumerate}
\label{pushout:assumptions}
    \item $Z$ is a \textit{descrete} e-hypergraph.
    \item $f,g$ are monos.
    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
    \item $\consistency^{\mu}(f(v_i)) = \consistency^{\mu}(f(v_j))$ and $\consistency^{\mu}(g(v_i)) = \consistency^{\mu}(g(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
\end{enumerate}    
then the pushout $X +_{f,g} Y$ exists.
\end{theorem}
\begin{proof}
    Consider the diagram below.
    \[
        \begin{tikzcd}
        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
        Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
                                                                        &                                                                  & Q
        \end{tikzcd}
    \]
    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
    First, a coproduct of $X+Y$ is computed which is 
    \[
        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{c,X+Y} \}
    \]
    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{c,X+Y}, \consistency_{X+Y}$.
    We will omit labels as they are irrelevant to pushout construction.
    Consider relations
    \begin{align*}    
    S_{V} &= \{\\
          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
          &\}
    \end{align*}
    \begin{align*}
    S_{E} &= \{\\
    &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
    &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
    &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
    &\}
    \end{align*}
and let relations $R_{V}$ and $R_{E}$ be their reflexive symmetric and transitive closures respectively.

@@

>>>>

CHRIS: REMARK THAT WE CANNOT DEAL WITH SIGNATURES WITH $f:0 \to 0$ MORPHISMS, AND WHY!
Because extended cospans have a more general notion of interface, DPOI rewriting as presented in Section \ref{sec:combinatorial-semantics} needs some adjustments.
First, the existence of \textit{internal} interfaces needs to be accounted for,  and in particular that internal interfaces are not necessarily preserved during rewriting.
The former means that we wish a rewrite rule to be a pair of \textit{extended} cospans of e-hypergraphs with matching external interfaces.
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\qquad \text{and} \qquad
    i \xrightarrow{} i'' \xrightarrow{} \mathcal{R} \xleftarrow{} j'' \xleftarrow{} j
\]
We will next use the fact that cospans
\[
    i \xrightarrow{} i' \xrightarrow{} \mathcal{L} \xleftarrow{} j' \xleftarrow{} j
\qquad \text{and} \qquad
    0 \xrightarrow{} 0 \xrightarrow{} \mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j
\]
endcode the same data and only consider cospans of the form
\[
\mathcal{L} \xleftarrow{} i' + j' \xleftarrow{} i + j \xrightarrow{} i'' + j'' \xrightarrow{} \mathcal{R}
\] to fit the DPO formalism.
We will futher omit the arrows $0 \to 0 \to \mathcal{L}$.
To illustrate the non-preservation of internal interfaces, consider the rewrite rule above and the following e-hypergraph with interfaces to perform the rewrite on
% \[\langle x \xrightarrow{} x' \xrightarrow{} \mathcal{L} \xleftarrow{} y' \xleftarrow{} y, x \xrightarrow{} x'' \xrightarrow{} \mathcal{R} \xleftarrow{} y'' \xleftarrow{} y \rangle \]
% such that $x' \not \cong x''$ and $y' \not \cong y''$ and consider the following e-hypergraph to perform the rewrite on
\[
    \mathcal{G} \xleftarrow{} n' + k' \xleftarrow{} n + k
\] such that there is a matching $m : \mathcal{L} \to \mathcal{G}$  and assume $i' \not \cong i''$ and $j' \not \cong j''$.
When removing the occurence of $\mathcal{L}$ from $\mathcal{G}$, \emph{i.e.}, by computing the pushout complement, the internal interfaces of $\mathcal{G}$ should be modified to exclude the vertices that were a part of (the image of) $\mathcal{L}$'s strict internal interfaces.
Then, when gluing $\mathcal{R}$ into the created context, we need to introduce strict internal interfaces of $\mathcal{R}$ into this context.
In particular, it means that the internal interfaces of the resulting cospan of e-hypergraphs
\[
    \mathcal{H} \xleftarrow{} n'' + k'' \xleftarrow{} n + k 
\]
not necessarily coincides with the internal interfaces of the initial cospan.
% The input interface on the left-hand side of the rule is $\{1, 2\}$, while the input interface on the right-hand side of the rule is $\{1,2,5,6,7,8\}$.
% Note that the outermost interfaces are preserved.

@@

\begin{figure}
    \[
    \scalebox{0.55}{
    \tikzfig{combinatorial_semantics/f_plus_g_new}
    }
    \]
    \caption{$+$ of two morphisms in $\Ecospans$}
    \label{fig:A+B}
\end{figure}

@@

Let the function $\llbracket - \rrbracket$ extend in the obvious way to apply to equations (\textit{i.e.}, pairs of $\Sigma^+$-terms). 
% \update{below definition depends on whether we require the homomorphism to reflect the conflicts}
CHRIS: COMMENT THAT WE QUOTIENT TO PHRASE CATEGORICALLY OUR RESULTS. 
\begin{definition}[Quotient by rewrites]  
Given a set of equations $\mathcal{E}$,  we denote by $\WellTypedMdaEcospans/\mathcal{E}$ the category $\WellTypedMdaEcospans$ quotiented by the following relation. 
\[
	f \sim g \quad \text{if} \quad f \Rrightarrow^{+,*}_{\llbracket \mathcal E \rrbracket} g ~ . 
\]
\end{definition}
In particular,  let $\mathcal{S}$ be the set equations of Figure \ref{fig:string-equations},  \textit{i.e.},  the semilattice equations over $\Sigma^+$-terms. 
This quotient in fact turns the interpretation function into a proper functor $\llbracket - \rrbracket: \textbf{PROP}^+(\Sigma) \to \WellTypedMdaEcospans/{\mathcal{S}}$,  which can further be quotiented to account for extra equations $\mathcal E$ arising in symmetric monoidal theories,  as below. 
\[
	\llbracket - \rrbracket_{\mathcal E}: \textbf{SMT}^+(\Sigma, \mathcal E) \to \WellTypedMdaEcospans/{\mathcal{S,E}}
\]

Note that not all pushouts (or pushout complements) in $\textbf{MEHypI}(\Sigma)$ are guaranteed to exist: however,  pushouts (and pushout complements) for rewrites generated from $\Sigma$-equations and those generated from the semilattice equations on $\Sigma^+$-terms do exist.  
%Let $\mathcal E$ be an arbitrary set of $\Sigma$-equations. 
Thus we have the following result. 
\begin{proposition}[Soundness]
\label{prop:soundness}
The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
\end{proposition}
It immediately follows that $\WellTypedMdaEcospans/{\mathcal{S,E}}$ is a semilattice-enriched PROP,  and $\llbracket - \rrbracket_{\mathcal E}$ is the free semilattice-enriched functor from $\textbf{SMT}^+(\Sigma, \mathcal E)$ induced by the interpretation of $\Sigma$ as itself. 
Note that the symmetric monoidal equations are absorbed in the hypergraph representation,  but the semilattice equations are covered by the fact we have quotiented our representation by certain rewrites.

CHRIS: NOTE ALSO WEAK NORMAL FORM FOR $\Sigma^+$-TERMS?
The equations of Figure~\ref{fig:string-equations} -- excepting the commutativity equation -- can all be directed in an obvious way to form a terminating rewrite system,  which results in a cospan in \textit{weak normal form}. 
\begin{proposition}[Weak normal form]
\label{prop:wnormal_form}
For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
\[
	f = f_1 + \ldots + f_n
\]
 such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
\end{proposition}

@@

Unfolding the definitions,  we can spell out the previous theorem in terms of DPO-rewriting. 
\begin{corollary}[Completeness of rewriting]
For any $f, g \in \textbf{SMT}^{+}(\Sigma, \mathcal{E})$,  we have
\[
	f = g \quad \iff \quad \llbracket f \rrbracket \Rrightarrow^{+,*}_{\mathcal{S}, \mathcal{E}} \llbracket g \rrbracket~ . 
\]
 \end{corollary}
%
%\begin{theorem}[Theorem 35~\cite{bonchi_string_2022-2}]
%
%    $a \Rightarrow_{R} b$ iff $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$.
%    Where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + \textsf{Frob} \to \catname{Csp_{D}(Hyp_{\Sigma})}$ and $\Rrightarrow$ is a convex rewriting relation for morphisms in $\catname{Hyp_{\Sigma}}$.
%\end{theorem}
%
%\begin{proposition}
%    $\llangle {}^{\ulcorner} a {}^{\urcorner} \rrangle \Rrightarrow_{\llangle {}^\ulcorner R {}^\urcorner \rrangle} \llangle {}^{\ulcorner}b {}^{\urcorner} \rrangle$ iff $[\mathcal{J}(a)] \Rrightarrow_{[R]}^{+} [\mathcal{J}(b)]$.
%     Note that $a$ and $b$ are from $\textsf{PROP}(\Sigma)$.
%\end{proposition}
%
%\begin{theorem}
%    $a \Rightarrow_{\langle l, r \rangle}^{+} b$ in $\textsf{PROP}^{+}$ iff $[a] \Rrightarrow_{\langle [l],[r] \rangle} [b]$ in $\sfrac{\WellTypedMdaEcospans}{\mathcal{S}}$.
%\end{theorem}
CHRIS: CONCLUDING STATEMENT.  COMMENT ABOUT GENERALIZATION TO DIRECTED REWRITES.

@@

\begin{theorem}[Existence of pushouts in $\catname{EHyp_{\Sigma}}$]
\label{th:existence_of_pushouts}
Consider the following span in $\catname{EHyp_{\Sigma}}$
% https://q.uiver.app/#q=WzAsMyxbMCwwLCJaIl0sWzIsMCwiWCJdLFswLDIsIlkiXSxbMCwxLCJmIl0sWzAsMiwiZyIsMl1d
\[\begin{tikzcd}
	Z && X \\
	\\
	Y
	\arrow["f", from=1-1, to=1-3]
	\arrow["g"', from=1-1, to=3-1]
\end{tikzcd}\]
such that
\begin{enumerate}
\label{pushout:assumptions}
    \item $Z$ is a \textit{discrete} e-hypergraph.
    \item $f,g$ are monos.
    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_{i},v_{j}$ in $V_{Z}$.
    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and if $[g_{V}(v)) \not = \varnothing$ then $[f_{V}(v)) = \varnothing$.
    \item $\consistency^{\mu}(f(v_i)) = \consistency^{\mu}(f(v_j))$ and $\consistency^{\mu}(g(v_i)) = \consistency^{\mu}(g(v_j))$ for all $v_i,v_j$ in $V_{Z}$.
\end{enumerate}    
then the pushout $X +_{f,g} Y$ exists.
\end{theorem}
\begin{proof}
    Consider the diagram below.
    \[
        \begin{tikzcd}
        Z \arrow[r, "f"] \arrow[d, "g"']                                   & X \arrow[d, "\sfrac{\iota_1}{\sim R}"] \arrow[rdd, "j_1", bend left] &   \\
        Y \arrow[r, "\sfrac{\iota_2}{\sim R}"] \arrow[rrd, "j_2"', bend right] & \sfrac{X+Y}{\sim R} \arrow[rd, "u"]                              &   \\
                                                                        &                                                                  & Q
        \end{tikzcd}
    \]
    Then, the pushout of e-hypergraphs $X$ and $Y$ is computed in two steps.
    First, a coproduct of $X+Y$ is computed which is 
    \[
        X + Y = \{V_{X} + V_{Y}, E_{X} + E_{Y}, s_{X+Y}, t_{X+Y}, \consistency_{X+Y}, <_{c,X+Y} \}
    \]
    where $s_{X+Y} : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ which can be defined as a copairing $[s'_{X}, s'_{Y}] : E_{X} + E_{Y} \to (V_{X} + V_{Y})^{*}$ and $s'_{X} : E_{X} \to (V_{X} + V_{Y})^{*}$, $s'_{Y} : E_{Y} \to (V_{X} + V_{Y})^{*}$ defined as $s'_{X} = s_{X};\iota_{1,V}^{*}$ and $s'_{Y} = s_{Y};\iota_{2,V}^{*}$ where $\iota_1,\iota_2$ are corresponding coproduct injections, similarly for $t_{X+Y}, <_{c,X+Y}, \consistency_{X+Y}$.
    We will omit labels as they are irrelevant to pushout construction.
    Consider relations
    \begin{align*}    
    S_{V} &= \{\\
          &(x_i,y_j) \in (V_{X} + V_{Y}) \times (V_{X} + V_{Y})\; | \\
          &\exists z \in V_{Z} \; . \; x_i = f_{V};\iota_{1,V}(z) \text{ and } y_j = g_{V};\iota_{2,V}(z)\\
          &\text{ where $x_i \in V_{X}$ and $y_j \in V_{Y}$ }\\
          &\}
    \end{align*}
    \begin{align*}
    S_{E} &= \{\\
    &(x_i,y_j) \in (E_{X} + E_{Y}) \times (E_{X} + E_{Y})\; |\\
    &\exists z \in E_{Z} \; . \; x_i = f_{E};\iota_{1,E}(z) \text{ and } y_j = g_{E};\iota_{2,E}(z)\\
    &\text{ where $x_i \in E_{X}$ and $y_j \in E_{Y}$}\\
    &\}
    \end{align*}
and let relations $R_{V}$ and $R_{E}$ be their reflexive symmetric and transitive closures respectively.

@@

\begin{proposition}[Proposition~\ref{prop:soundness} restatement]
The category $\WellTypedMdaEcospans/{\mathcal{S}}$ is a semilattice-enriched PROP. 
\end{proposition}

\begin{proposition}[Proposition~\ref{prop:wnormal_form} restatement]

For each cospan $f$ in ${\WellTypedMdaEcospans}/{\mathcal{S}}$ there is a \textit{weak normal form} such that 
\[
    f = f_1 + \ldots + f_n
\]
such that each $f_i$ contains no hierarchical edges,  and for all $i \neq j$ we have $f_i \neq f_j$.
\end{proposition}
